\chapter{The General Two-Higgs Doublet Model}
\label{ch:g2HDM}
There have been a plethora of 2HDMs proposed throughout the years, each model with its own set of additional assumptions.
A comprehensive review of various 2HDMs can be found in~\cite{Branco20122HDMs}.

\section{The Model}
In the "General Two Higgs Doublet Model", or {\gthdm} for short,
we introduce a second Higgs doublet without imposing any \(Z_{2} \) symmetry. 
Having no extra symmetry assumptions also means the fermions can couple to both Higgs doublets.
In that case, the most general Yukawa interaction for up-type quarks can be written as
\begin{equation}\label{eq:two_higgs_lagrangian}
  \mathcal{L} = -\bar{q}_{iL}(Y_{1ij}^{u}\tilde{\Phi_{1}} + Y_{2ij}^{u}\tilde{\Phi_{2}})u_{jR} + \mathrm{h.c.}
\end{equation}
where \(i \text{, } j \) are flavor indices, \(\tilde{\Phi_{b}} = i\sigma_{2}\Phi_{b}^{*}\,(b = 1 \text{, } 2) \) with
\begin{equation}
  \Phi_{b}(x) = 
  \begin{pmatrix}
    \phi_{b}^{+} \\
    \frac{1}{\sqrt{2}(v_{b}+h_{b}(x)+ia_{b}(x))}
  \end{pmatrix},
\end{equation}
\(\sigma_{2} \) being a Pauli matrix.
We can denote thte vevs as \(v_{1} = v c_{\beta} \) and \(v_{2} = v s_{\beta} \), \(v = \SI{246}{\GeV} \).
We will be using the shorthand notation \(s_{\beta} \equiv \sin{\beta} \), \(c_{\beta} \equiv \cos{\beta} \), and \(t_{\beta} \equiv \tan{\beta} \).
We usually prefer to use the basis in which only one of the two Higgs doublets has vev.
Under this basis, the {\gthdm} Lagrangian can be written as~\cite{DavidsonHaber20052HDM,HouModak2020G2HDM}
\begin{adjustwidth}{-1.5cm}{-1.5cm}
  \begin{align}
    \mathcal{L} = - & \frac{1}{\sqrt{2}} \sum_{f = u, d, \ell} \bar f_{i} \Big[\big(-\lambda^f_i \delta_{ij} s_\gamma + \rho^f_{ij} c_\gamma\big) h
    + \big(\lambda^f_i \delta_{ij} c_\gamma + \rho^f_{ij} s_\gamma\big)H
    - i\,{\rm sgn}(Q_f) \rho^f_{ij} A\Big]  R\, f_{j} \nonumber \\
    - & \bar{u}_i\left[(V\rho^d)_{ij} R-(\rho^{u\dagger}V)_{ij} L\right]d_j H^+
    - \bar{\nu}_i\rho^L_{ij} R \, \ell_j H^+ + {\rm h.c.},
    \label{eq:g2hdm_lagrangian}
  \end{align}
\end{adjustwidth}
where the generation indices \(i \), \(j \) are summed over, \(L \), \(R = (1\pm\gamma_{5})/2\) are projections, \(V \) is the CKM matrix for quarks and unity for leptons.
\(\lambda^f \) are the SM Yukawa matrices, and \(\rho^f \) are the extra-Yukawa matrices.
These two matrices are related to the \(Y_{1,2} \) in~\eqnref{eq:two_higgs_lagrangian} by
\begin{equation}\label{eq:y_and_lambda_rho_relation}
  \lambda = V_{L}^{u\dagger}(Y_{1}c_{\beta} + Y_{2}s_{\beta})V_{R}^{u}\, \text{, } \quad 
  \rho = V_{L}^{u\dagger}(-Y_{1}s_{\beta} + Y_{2}c_{\beta})V_{R}^{u}
\end{equation}
A key takeaway is that each family of fermions (u-type, d-type, lepton) is associated with its own extra-Yukawa \(\rho \) matrix.
Having no imposed discrete symmetry imposed does have its consequences;
namely, they give rise to possibly large flavor-changing neutral Higgs (FCNH) couplings.
These couplings are essentialy the off-diagonal terms of the \(\rho \) matrices.
In {\gthdm}, these FCNH couplings are controlled by \textit{flavor hierarchies} and \textit{alignment}.
The gist of \emph{flavor hierarchies} means that the \(\rho \) matrices somehow ``know'' the current flavor structure of the SM, 
namely the mass hierarchy, but also the CKM matrix.
Cheng and Sher paved the way with their ansatz~\cite{ChengSher1987Ansatz} of \(\rho_{ij} \sim \sqrt{m_{i}m_{j}}\) early on,
and in {\gthdm} this concept has been generalized to the ``rule of thumb''~\cite{HouKumar2020RuleOfThumb}
\begin{equation}
  \rho_{ii} \lesssim \order{\lambda_i}, \quad
  \rho_{1i} \lesssim \order{\lambda_1}, \quad
  \rho_{3j} \lesssim \order{\lambda_3},
  \label{eq:ruleofthumb}
\end{equation}
with \(j\neq 1 \).
The sheer discrepancy of the mass scale between the top quark and everything else serves as a layer of protection,
keeping most of the FCNH \(\rho \) couplings small, 
while still providing adequate parameter space to explore for NP in the top-related couplings.
The second controlling mechanism \emph{alignment} emerged from the experimental discovery of the \(h(125) \) boson: 
the \(h \) boson in the Lagrangian closely resembles the SM \(h(125) \).
This in turn means that the \(h-H \) mixing angle \(c_{\gamma} \) is small~\cite{HouKikuchi2018Alignment}.
Consequently, the SM-like Higgs \(h \) is mostly controlled by the SM Yukawas, while the newly introduced \(\rho \) matrices control the exotic Higgses \(H, A, H^{\pm} \).
This extra \(c_{\gamma} \) suppression provides an additional layer of protection against the FCNH \(\rho \) couplings.
A remarkable feature of {\gthdm} is that \(\order{1}\,\rho_{tt}\) can drive EWBG through~\cite{FHS2018EWBGandEDM} \(\lambda_{t}\Im\rho_{tt} \).

\section{Electroweak Baryogenesis}\label{sec:EWBG}
Even though the SM Lagrangian is baryon number conserving, 
there still is the possibility of baryon-number-violating processes due to a quantum anomaly~\cite{tHooft1976BaryonNumAnomaly}.
This anomaly gives rise to baryon-number-violating but \((B-L) \) conserving sphaleron transitions~\cite{Klinkhamer1984Sphaleron}, where \(B \) is baryon number and \(L \) is lepton number.
The rate of these transitions \(\Gamma_{B} \) scale with temperature as \(\sim \exp\left[-\frac{\pi m_{W}(T)}{\alpha_{W}}\right] \)~\cite{Kuzmin1985EWBG},
with \(m_{W} \) the W boson mass and \(\alpha_{W} \) the weak coupling constant,
which means that said transitions are more rapid in the earlier symmetric phase of the universe,
before the symmetry breaking phase transition that induces the Higgs vev occurs.
If \(\Gamma_{B} \) is larger than the expansion rate of the universe, namely the Hubble parameter \(H \),
then any ``primordial'' baryon asymmetry will be washed out by these sphaleron processes.
This implies that the EWPT must be a first-order phase transition, 
and that the vev aquired in the broken phase must be large enough such that \(\Gamma_{B}^{(\mathrm{br})}(T_{C}) < H(T_{C})\) at critical temperature \(T_{C} \).
This ensures that any baryon asymmetry generated by \emph{the phase transition itself} will persist without being washed out.
During this phase transition, ``bubbles'' of the new broken phase nucleate and expand to fill the universe~\cite{Farrar1993MSMBAU}.
This creates a departure from thermal equilibrium, satisfying the third Sakharov condition.
If there exists CPV in the particle interactions that happen at the bubble walls,
then BAU can be generated.
Unfortunately, as mentioned before, the CPV present in the SM is insufficient to drive the BAU of our current observable universe.

In {\gthdm}, we have an extra Higgs doublet, as well as extra Yukawa couplings to work with.
Thermal loops of heavy Higgs bosons can create a strong enough first-order EWPT~\cite{Bochkarev1990EWBG2HDM} to satisfy the above criteria,
doing to \(\order{1} \) nondecoupled~\cite{Kanemura2005EWBGHiggs} Higgs couplings.
We can estimate BAU by~\cite{HuetNelson1996EWBG, ClineKainulainen2000EWBG}
\begin{equation}
  Y_{B} \equiv \frac{n_{B}}{s} = \frac{-3\Gamma_{B}^{(\mathrm{sym})}}{2D_{q}\lambda_{+}s}\int_{-\infty}^{0}\dd{z^{\prime}}n_{L}(z^{\prime})e^{-\lambda_{-}z^{\prime}}
\end{equation}
where \(D_{q} \simeq 8.9/T \) is the quark diffusion constant, \(s \) is the entropy density,
\(\Gamma_{B}^{\mathrm{sym}} = 120\alpha_{W}^{5}T \) is the \(B \)-changing rate in the symmetric phase,
and \(\lambda_{\pm} = [v_{w} \pm (v_{w}^{2} + 15\Gamma_{B}^{\mathrm{sym}}D_{q})^{1/2}]/2D_{q} \),
with \(v_{w} \) the bubble wall velocity.
The integration happens over \(z^{\prime} \), which is the coordinate opposite of the bubble wall expansion direction.
For \(Y_{B} \) to be nonzero, a nonvanishing total left-handed fermion number density \(n_{L} \) is needed.
The current observational value \(Y_{B}^{\mathrm{obs}} = \num{8.59e-11} \) comes from the Planck collaboration~\cite{Planck2014Yobs}.
The BAU-related CPV arises from the interaction between particles/antiparticles with the bubble wall.
As a result of their thermal motion, particles and antiparticles in the neighborhood of the bubble wall propagate through it. 
Since one side of the bubble wall has zero vev, while the other side has nonzero vev,
the particles see the bubble wall as a potential barrier and scatter from it.
A dominant process is shown in~\figref{fig:BAU-CPV_process}.
The Higgs bubble wall is denoted as \(v_{a}(x)\text{, }v_{b}(y) \), which are spacetime-dependent vevs.
The CPV source term \(S_{ij} \) for left-handed fermion \(f_{iL} \) induced by right-handed fermion \(f_{jR} \) in {\gthdm} is~\cite{FHS2018EWBGandEDM}
\begin{equation}
  S_{iLjR}(Z) = N_{C}F\Im[(Y_{1})_{ij}(Y_{2})_{ij}^{*}]v^{2}(Z)\partial_{t_{Z}}\beta(Z)
\end{equation}
where \(Z = (t_{Z}, 0, 0, z) \) is the position in the heat bath, \(N_{C} = 3 \) is the number of color,
and \(F \) is a function (see Ref.~\cite{CFS2016EWBG}).
Essentially, the CPV for BAU is contained in \(\Im[(Y_{1})_{ij}(Y_{2})_{ij}^{*}] \).
Using~\eqnref{eq:y_and_lambda_rho_relation}, 
assuming \((Y_{1,2})_{ij} = 0 \) except for \((Y_{1,2})_{tc} \) as well as \((Y_{1})_{tt} = (Y_{2})_{tt} \) with \(t_{\beta} = 1 \),
solving for \(V_{R}^{u} \) in terms of \(Y_{1,2} \) gives
\begin{equation}
  \Im[(Y_{1})_{tc}(Y_{2})_{tc}^{*}] = -\lambda_{t}\Im\rho_{tt}\, \text{,} \quad \rho_{tc} = 0
\end{equation}
And thus we arrive at the most important result of~\cite{FHS2018EWBGandEDM}: BAU can be driven in the {\gthdm} by \(\lambda_{t}\Im\rho_{tt} \).

\clearpage
\input{body/2_framework/g2HDM_figures.tex}